---
title: "Untitled"
author: "Maxime Dahirel"
date: "19/05/2020"
output: 
  html_document:
    theme: yeti
    toc: TRUE
    toc_float: TRUE
    code_download: TRUE
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)

library(coda)
library(rstan)
library(bayesplot)
library(brms)
rstan_options(auto_write = TRUE)
#options(mc.cores = 2) #temporary unparallel to view progress in console rather than viewer to avoid viewer bug in new rstudio preview
library(brms)

library(tidybayes)


library(patchwork)

library(here)
```

# Context


#load data

```{r}

raw_dynamics <- read_csv(here("data","expansion_data","Trichogramma_dynamics.csv"))

raw_size <- read_csv(here("data","exp1_bodysize.csv"))

raw_mvt <- read_csv(here("data","exp2_movement.csv"))

raw_fec <- read_csv(here("data","exp2_fecundity.csv"))
raw_disp <- read_csv(here("data","exp2_dispersal.csv"))

raw_fec_dens <- read_csv(here("data","exp3_fecundity.csv"))
raw_disp_dens <- read_csv(here("data","exp3_dispersal.csv"))
```

#Population density (reprisal)

```{r}
data_popsize<- raw_dynamics %>% 
  mutate(Mix= as.numeric(str_sub(Image,1,1)),
         Treatment = str_sub(Image,2,3),
         Replicate = as.numeric(str_sub(Image,4,4)),
         Patch = as.numeric(str_sub(Image,6,7))
  ) %>% 
  #no further processing needed; we exploit the fact that chars 6&7 are either digit dot (patchs 0 to 9)
#or digit-digit (patches 10 and beyond)
## as.numeric() resolves both correctly (e.g; "2." and "12" become 2 and 12)
  mutate(landscapeID=paste("Mix",Mix,"_Treatment",Treatment,"_Replicate",Replicate,sep="")) %>% #a unique replicate ID
  mutate(Peggs_est=P/(P+H), ## proportion of pixels counted as parasitised (estimated)
         spacetimeID=paste(landscapeID,"_Generation",Generation,"_Patch",Patch,sep=""), # a unique ID for each replicate x patch x generation combination
         Mix = factor(Mix)) %>% 
  mutate(pixtot=P+H) %>% 
  mutate(Treatment = fct_recode(Treatment,`control`="PL",`restricted connectedness`="PS")) %>% 
  group_by(landscapeID,Generation) %>% 
  mutate(front=max(Patch)) %>% 
  ungroup() %>% 
  group_by(landscapeID,Patch) %>% 
  mutate(founding_time=min(Generation)) %>% 
  ungroup() %>% 
  mutate(distance_from_front=front-Patch,
         time_since_founding=Generation-founding_time) %>% 
  select(Macro,landscapeID,Patch,Generation,Treatment,distance_from_front,Peggs_est,Mix,front,pixtot,time_since_founding,founding_time)

front_prev<-data_popsize %>% 
  select(landscapeID,Generation,front) %>% 
  distinct() %>% 
  group_by(landscapeID) %>%
  arrange(Generation) %>% 
  mutate(front_prev=lag(front)) %>% 
  mutate(front_prev=replace_na(front_prev,0)) 

data_popsize<-data_popsize %>% filter(Patch>0) %>% 
  select(landscapeID,Patch,Generation) %>% 
  distinct() %>% #remove macro induced quadruplicates
  left_join(.,front_prev) %>% 
  group_by(landscapeID,Patch) %>% 
  filter(Generation==min(Generation)) %>% 
  mutate(dist_to_front_prev=Patch-front_prev) %>% #<0 in few cases; patches "passed over" during colonisation, then gap colonised later
  ungroup() %>%
  select(landscapeID,Patch,dist_to_front_prev) %>% 
  left_join(data_popsize,.) %>% filter(Patch>0) %>% 
  mutate(scale_Generation=scale(Generation)) %>% 
  mutate(spacetimeID=paste(landscapeID,Generation,Patch)) %>% 
  ungroup()

##to do : how to manage extinct-then-recolonised patches?

```

```{r}
#monomolecular: y = K-(K-M0)*exp(-r*distance_from_front)
#3par logistic y = M0*K/(M0+(K-M0)*exp(-r*distance_from_front))
#gompertz y = K*(M0/K)^(exp(-r*distance_from_front))
#taken from paine et al 2012 MEE

mod_popsize <- brm(bf(Peggs_est~logit(K-(K-M0)*exp(-r*distance_from_front))+obs,
                      nlf(K~inv_logit(logitK)),
                      nlf(M0~inv_logit(logitM0)),
                      nlf(r~10^log10r),
                      logitK~scale_Generation*Treatment+(1|Mix)+(1|Macro)+(scale_Generation|1|landscapeID),
                      logitM0~(scale_Generation*Treatment):new_front+(1|Mix)+(1|Macro)+(scale_Generation*new_front|1|landscapeID),
                      log10r~(scale_Generation*Treatment):new_front+(1|Mix)+(1|Macro)+(scale_Generation*new_front|1|landscapeID),
                      nlf(phi~10^log10phi),
                      log10phi~(1|Macro),
                      obs~0+(1|Macro)+(1|spacetimeID),
                      nl=TRUE),
        data=data_popsize,family=Beta(link_phi="identity"), #careful with link_phi = identity, avoid if covariates, especially ranefs; but think how to setup otherwise
        iter=350,warmup=150,chains=2,
        prior=c(
          set_prior("normal(0,1.5)",class="b",nlpar=c("logitK","logitM0","log10r")),
          set_prior("normal(0,1)",class="sd",nlpar=c("logitK","logitM0","log10r")),
          set_prior("normal(0,1)",class="sd",nlpar="obs"),
          set_prior("normal(0,1.5)",class="b",nlpar="log10phi"),
          set_prior("normal(0,1)",class="sd",nlpar="log10phi")
        ),
        control=list(adapt_delta=0.8),seed=42)
```

testing something 
try the logit(y)~logit(formula),family=student trick again?
try phi being nonlinear the same way asmu?

```{r}

mod_popsize <- brm(bf(Peggs_est~logit(K-(K-M0)*exp(-r*time_since_founding))+obs,
                      nlf(K~inv_logit(logitK)),
                      nlf(M0~inv_logit(logitM0)),
                      nlf(r~10^log10r),
                      logitK~0+Treatment+(1|Mix)+(1|1|landscapeID)+(1|2|landscapeID:Patch),
                      logitM0~0+Treatment*scale(dist_to_front_prev)+(1|Mix)+(1|1|landscapeID)+(scale(dist_to_front_prev)|2|landscapeID:Patch),
                      log10r~0+Treatment*scale(dist_to_front_prev)+(1|Mix)+(1|1|landscapeID)+(scale(dist_to_front_prev)|2|landscapeID:Patch),obs~0+(1|Macro)+(1|spacetimeID),
                      nlf(phi~10^log10phi),
                      log10phi~1,
                      nl=TRUE),
                   data=data_popsize,family=Beta(link_phi="identity"), #careful with link_phi = identity, avoid if covariates, especially ranefs; but think how to setup otherwise
                   iter=200,warmup=100,chains=2,
                   prior=c(
                       set_prior("normal(0,1.5)",class="b",nlpar=c("logitK","logitM0","log10r")),
                       set_prior("normal(0,1)",class="sd",nlpar=c("logitK","logitM0","log10r","obs")),
                       set_prior("normal(0,1.5)",class="b",nlpar="log10phi")
                   ),
                   control=list(adapt_delta=0.8),seed=42)
```


```{r}
data_popsize2<-data_popsize %>% 
  filter(Patch ==front | Patch==0) %>% 
  mutate(Location=ifelse(Patch==0,"core",paste("front",new_front))) %>% 
  mutate(Location=fct_recode(factor(Location),`new_front`="front TRUE",`old_front`="front FALSE"))
```

```{r}
#linear version

mod_popsize <- brm(bf(Peggs_est~0+Location:Treatment+Location:Treatment:scale_Generation+(1|Macro)+(1|Mix/landscapeID/spacetimeID),
                      nlf(phi~1/invphi),
                      invphi~1),
        data=data_popsize2,family=Beta(link_phi="identity"), #careful with link_phi = identity, avoid if covariates, especially ranefs; but think how to setup otherwise
        iter=350,warmup=150,chains=2,
        prior=c(
          set_prior("normal(0,1.5)",class="b"),
          set_prior("normal(0,1)",class="sd"),
          set_prior("normal(0,1)",nlpar="invphi",lb=0)
          ),
        control=list(adapt_delta=0.8),seed=42)
```

â•not much new compared to suppl of 1st manuscript. Not mention it, send reader back to 1st manuscript to justify higher pop sizes in restr conn? (what about diff core/front to justify the experimental low/high? we can do without at least in first round, not a perfect match anyway)

```{r}
data_popsize %>% ungroup() %>% 
  mutate(Macro=Macro[1],landscapeID=landscapeID[1],Generation=Generation[1]) %>% 
  select(Macro,landscapeID,Generation,Treatment) %>% 
  distinct() %>% 
  expand_grid(distance_from_front=c(0:20)) %>% 
  add_fitted_draws(mod_popsize,re_formula = NA) %>% 
  ggplot() +
  stat_lineribbon(aes(x=distance_from_front,y=.value))+facet_wrap(~Treatment)
```


# Body size

```{r}
data_size<-raw_size %>%
                mutate(IDgroup=ifelse(Generation==0,
                                      paste("Mix",Mix,"_stock",sep=""),
                                      paste("Mix",Mix,"_",Treatment,"_Rep",Replicate,"_GenerationFinal",sep="")
                                      )
                       )%>% 
                mutate(IDpop=ifelse(Generation==0,
                                    IDgroup,
                                    paste(IDgroup,"_Location",Location,sep=""))) %>% 
                mutate(IDindiv=paste(IDpop,ID_in_batch,sep="_")) %>% 
                mutate(trt=paste(Treatment,Location)) %>% 
                mutate(trt=relevel(factor(trt),"stock stock")) %>%
                mutate(trt=fct_recode(trt,stock="stock stock")) %>% 
                pivot_longer(cols=c("tibia_obsA","tibia_obsC"),names_to="observer",values_to="tibia") %>% 
                mutate(is.obsA=-0.5+1*(observer=="tibia_obsA"))

```

use of nested hieracrhic groups to reflect phylogenetic relationships among pop taken from 

```{r}
mod_size=brm(bf(scale(tibia)~0+trt+is.obsA+(1|Mix/IDgroup/IDpop)+(1|gr(IDindiv,by=trt))),
        data=data_size,
        chains=4,iter=2000,
        prior=c(set_prior("normal(0,1)",class="b"),
                set_prior("normal(0,1)",class="sigma"),
                set_prior("normal(0,1)",class="sd")),
        seed=42
)

```

shifts mostly in variance. Looks negatively correlated with population density on aevrage in each category (though no way to be sure without stock densities)
read on effect of competition on size both plastic and selective


find a way to model the among-pop variance as one main variance pop + 1 dirichelt something to split in mix, group, pop???
possible??
# Activity

```{r}
data_mvt<-raw_mvt %>% 
                filter(Generation %in% c(0,10)) %>% 
                filter(timestamp<=300) %>% ##some replicate are filmed longer than the nominal 5 minutes, this removes their ends to standardize
                mutate(Nobs=substr(File,nchar(File)-1,nchar(File))) %>% #note: rerecord correctly #wrong for G4 for which not given and last code is testing unit
                mutate(Nobs=as.numeric(str_remove(Nobs,"_"))) %>%  ##but we don't do G4 anyway, soooo
                mutate(IDgroup=ifelse(Generation==0,
                                      paste("Mix",Mix,"_stock",sep=""),
                                      paste("Mix",Mix,"_",Treatment,"_Rep",Replicate,"_GenerationFinal",sep="")
                                      )
                       )%>% 
                mutate(IDpop=ifelse(Generation==0,
                                    IDgroup,
                                    paste(IDgroup,"_Location",Location,sep=""))) %>% 
                mutate(IDtest=paste(IDpop,testing_unit,sep="_")) %>% 
                mutate(trt=paste(Treatment,Location)) %>% 
                mutate(trt=relevel(factor(trt),"stock stock")) %>%
                mutate(trt=fct_recode(trt,stock="stock stock")) %>% 
                group_by(Mix,Treatment,Generation,Replicate,Location,IDgroup,IDpop,IDtest,trt) %>% 
                summarise(percentData=mean(percentData), ##this column is already a testing_unit- level variable
                          activity_time=sum(path_duration*activity_percent),
                          obs_time=sum(path_duration),
                          Nobs=mean(Nobs)) %>% 
                mutate(activity_percent=activity_time/obs_time) %>% 
  ungroup()
```


```{r}
mod_mvt=brm(bf(activity_percent~0+trt+(1|Mix/IDgroup/IDpop),
               nlf(phi~10^log10phi),
               log10phi~percentData),
        data=data_mvt,family=Beta,
        chains=4,iter=2000,
        prior=c(set_prior("normal(0,1.5)",class="b"),
                set_prior("normal(0,1)",class="sd"),
                set_prior("normal(0,1)",class="b",nlpar="log10phi")),
        seed=42
)

```

# Fecundity 1

```{r}
data_fec<-raw_fec %>% 
                filter(Generation %in% c(0,12)) %>%
                mutate(IDgroup=ifelse(Generation==0,
                                      paste("Mix",Mix,"_stock",sep=""),
                                      paste("Mix",Mix,"_",Treatment,"_Rep",Replicate,"_GenerationFinal",sep="")
                                      )
                       )%>% 
                mutate(IDpop=ifelse(Generation==0,
                                    IDgroup,
                                    paste(IDgroup,"_Location",Location,sep=""))) %>% 
                mutate(trt=paste(Treatment,Location)) %>% 
                mutate(trt=relevel(factor(trt),"stock stock")) %>%
                mutate(trt=fct_recode(trt,stock="stock stock"))
```


```{r}
mod_fec=brm(bf(Fecundity~0+trt+(1|Mix/IDgroup/IDpop),
            zi~0+trt+(1|Mix/IDgroup/IDpop),
            nlf(shape~1/invshape),invshape~1),
         data=data_fec,family=zero_inflated_negbinomial(link_shape="identity"), #or hurdle?? think about it
         chains=4,iter=2000,
         prior=c(set_prior("normal(0,1.5)",class="b"),
                 set_prior("normal(0,1)",class="sd"),
                 set_prior("normal(0,1.5)",class="b",dpar="zi"),
                 set_prior("normal(0,1)",class="sd",dpar="zi"),
                 set_prior("normal(0,1)",nlpar="invshape",lb=0))
)
```

# Dispersal 1
```{r}
data_disp<-raw_disp %>% 
                filter(Generation %in% c(0,12)) %>%
                mutate(IDgroup=ifelse(Generation==0,
                                      paste("Mix",Mix,"_stock",sep=""),
                                      paste("Mix",Mix,"_",Treatment,"_Rep",Replicate,"_GenerationFinal",sep="")
                                      )
                       )%>% 
                mutate(IDpop=ifelse(Generation==0,
                                    IDgroup,
                                    paste(IDgroup,"_Location",Location,sep=""))) %>% 
                mutate(trt=paste(Treatment,Location)) %>% 
                mutate(trt=relevel(factor(trt),"stock stock")) %>%
                mutate(trt=fct_recode(trt,stock="stock stock")) %>% 
                mutate(Neggs_all=Neggs_start+Neggs_arrival) %>% 
                mutate(Neggs_all_scaled=scale(Neggs_all))
```


```{r}
mod_disp=brm(bf(Neggs_arrival|trials(Neggs_all)~0+trt+Neggs_all_scaled+(1|Mix/IDgroup/IDpop)),
         data=data_disp,family=binomial,
         chains=4,iter=200,
         prior=c(set_prior("normal(0,1.5)",class="b"),
                 set_prior("normal(0,1)",class="sd"))
)
```

loo/kfold to compare to version with no Neggs_scaled? seems to work (both here and density) but need to recheck because pareto bad obs

```{r}
mod_disp_fec=brm(bf(Neggs_all~0+trt+(1|Mix/IDgroup/IDpop)),
         data=data_disp,family=negbinomial,
         chains=4,iter=200,
         prior=c(set_prior("normal(0,1.5)",class="b"),
                 set_prior("normal(0,1)",class="sd"))
)
```


# Fecundity 2
```{r}
data_fec_dens<-raw_fec_dens %>% 
                mutate(IDgroup=ifelse(Generation==0,
                                      paste("Mix",Mix,"_stock",sep=""),
                                      paste("Mix",Mix,"_",Treatment,"_Rep",Replicate,"_GenerationFinal",sep="")
                                      )
                       )%>% 
                mutate(IDpop=ifelse(Generation==0,
                                    IDgroup,
                                    paste(IDgroup,"_Location",Location,sep=""))) %>% 
                mutate(trt=paste(Treatment,Location)) %>% 
                mutate(trt=relevel(factor(trt),"stock stock")) %>%
                mutate(trt=fct_recode(trt,stock="stock stock"))%>% 
                mutate(Density_centred=-0.5+1*(Density=="high"))
```


```{r}
mod_fec_dens=brm(bf(Fecundity~0+trt*Density_centred+(1|Mix/IDgroup/IDpop),
            zi~0+trt*Density_centred+(1|Mix/IDgroup/IDpop),
            nlf(shape~1/invshape),invshape~1),
         data=data_fec_dens,family=zero_inflated_negbinomial(link_shape = "identity"),
         chains=4,iter=200,
         prior=c(set_prior("normal(0,1.5)",class="b"),
                 set_prior("normal(0,1)",class="sd"),
                 set_prior("normal(0,1.5)",class="b",dpar="zi"),
                 set_prior("normal(0,1)",class="sd",dpar="zi"),
                 set_prior("normal(0,1)",nlpar="invshape",lb=0))
)
```
# Dispersal 2

```{r}
data_disp_dens<-raw_disp_dens %>% 
                mutate(IDgroup=ifelse(Generation==0,
                                      paste("Mix",Mix,"_stock",sep=""),
                                      paste("Mix",Mix,"_",Treatment,"_Rep",Replicate,"_GenerationFinal",sep="")
                                      )
                       )%>% 
                mutate(IDpop=ifelse(Generation==0,
                                    IDgroup,
                                    paste(IDgroup,"_Location",Location,sep=""))) %>% 
                mutate(trt=paste(Treatment,Location)) %>% 
                mutate(trt=relevel(factor(trt),"stock stock")) %>%
                mutate(trt=fct_recode(trt,stock="stock stock")) %>% 
                mutate(Neggs_all=Neggs_start+Neggs_arrival) %>% 
                mutate(Neggs_all_scaled=scale(Neggs_all)) %>% 
                mutate(Neggs_all_log=log(Neggs_all)) %>% 
                mutate(Density_centred=-0.5+1*(Density=="high"))
```


```{r}
mod_disp_dens=brm(bf(Neggs_arrival|trials(Neggs_all)~0+trt*Density_centred+Neggs_all_scaled+(1|Mix/IDgroup/IDpop)),
         data=data_disp_dens,family=binomial,
         chains=4,iter=200,
         prior=c(set_prior("normal(0,1.5)",class="b"),
                 set_prior("normal(0,1)",class="sd"))
)
```


```{r}
mod_disp_dens_fec=brm(bf(Neggs_all~0+trt*Density_centred+(1|Mix/IDgroup/IDpop)),
         data=data_disp_dens,family=negbinomial,
         chains=4,iter=200,
         prior=c(set_prior("normal(0,1.5)",class="b"),
                 set_prior("normal(0,1)",class="sd"))
)
```