---
title: "Untitled"
author: "Maxime Dahirel"
date: "19/05/2020"
output: 
  html_document:
    theme: yeti
    toc: TRUE
    toc_float: TRUE
    code_download: TRUE
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)

library(rstan)
library(bayesplot)
library(brms)
rstan_options(auto_write = TRUE)
#options(mc.cores = 2) #temporary unparallel to view progress in console rather than viewer to avoid viewer bug in new rstudio preview
library(brms)

library(tidybayes)


library(patchwork)

library(here)
```

# Context


#load data

```{r}

raw_dynamics <- read_csv(here("data","expansion_data","Trichogramma_dynamics.csv"))

raw_size <- read_csv(here("data","exp1_bodysize.csv"))

raw_mvt <- read_csv(here("data","exp2_movement.csv"))

raw_fec <- read_csv(here("data","exp2_fecundity.csv"))
raw_disp <- read_csv(here("data","exp2_dispersal.csv"))

raw_fec_dens <- read_csv(here("data","exp3_fecundity.csv"))
raw_disp_dens <- read_csv(here("data","exp3_dispersal.csv"))
```

#Population density (reprisal)

```{r}
data_popsize<- raw_dynamics %>% 
  mutate(Mix= as.numeric(str_sub(Image,1,1)),
         Treatment = str_sub(Image,2,3),
         Replicate = as.numeric(str_sub(Image,4,4)),
         Patch = as.numeric(str_sub(Image,6,7))
  ) %>% 
  #no further processing needed; we exploit the fact that chars 6&7 are either digit dot (patchs 0 to 9)
#or digit-digit (patches 10 and beyond)
## as.numeric() resolves both correctly (e.g; "2." and "12" become 2 and 12)
  mutate(landscapeID=paste("Mix",Mix,"_Treatment",Treatment,"_Replicate",Replicate,sep="")) %>% #a unique replicate ID
  mutate(Peggs_est=P/(P+H), ## proportion of pixels counted as parasitised (estimated)
         obsID=paste(landscapeID,"_Generation",Generation,"_Patch",Patch,sep=""), # a unique ID for each replicate x patch x generation combination
         Mix = factor(Mix)) %>% 
  mutate(Treatment = fct_recode(Treatment,`reference`="PL",`restricted connectedness`="PS")) %>% 
  group_by(landscapeID,Generation) %>% 
  mutate(front=max(Patch)) %>% 
  ungroup() %>% 
  group_by(landscapeID,Patch) %>% 
  arrange(Generation) %>% 
  mutate(front_prev=ifelse(Generation==1,0,lag(front))) %>% 
  mutate(recol=Generation>(lag(Generation)+1)) %>% ##if there's a gap, it means extinction followed by recolonisation later
  mutate(recol=replace_na(recol,FALSE)) %>% 
  filter(Patch>0) %>%  ## we don't use the release patch
  filter(cumsum(recol)==0) %>%    ## we only keep initial colonisations, so remove everything after initial extinction in a patch
  mutate(founding_generation=min(Generation)) %>% 
  mutate(time_since_founding=Generation-founding_generation) %>% 
  
  #mutate(is.passed.over=sum(time_since_founding==0 & Patch<front_prev,na.rm=TRUE)) %>% 
  #an indicator to say if the patch was first founded from behind the previous front, instead of by the front advance
  #filter(is.passed.over==0) %>% 
  ungroup() %>%
  select(Macro,landscapeID,Patch,Generation,Treatment,obsID,Peggs_est,Mix,time_since_founding)

##to do : count local extinction and count "passed over"; maybe actually include them if too big ????

```



```{r}

#monomolecular: y = K-(K-F)*exp(-r*time_since_founding)
#3par logistic y = F*K/(F+(K-F)*exp(-r*time_since_founding))
#gompertz y = K*(F/K)^(exp(-r*time_since_founding))
#taken from paine et al 2012 MEE

mod_popsize <- brm(bf(Peggs_est~unobservedtrue+obsprocess,
                      nlf(unobservedtrue~trend+residuals),
                      nlf(trend~logit(K-(K-F)*exp(-r*time_since_founding))),
                      nlf(K~inv_logit(logitK)),
                      nlf(F~inv_logit(logitF)),
                      nlf(r~exp(logr)),
                      logitK~0+Treatment+(1|Mix)+(1|1|landscapeID)+(1|2|landscapeID:Patch),
                      logitF~0+Treatment+(1|Mix)+(1|1|landscapeID)+(1|2|landscapeID:Patch),
                      logr~0+Treatment+(1|Mix)+(1|1|landscapeID)+(1|2|landscapeID:Patch),
                      nlf(phi~1/invphi),
                      invphi~1,
                      residuals~0+(1|gr(obsID,by=Treatment)),
                      obsprocess~0+(1|Macro),
                      nl=TRUE),
                   data=data_popsize,family=Beta(link_phi="identity"), #careful with link_phi = identity, avoid if covariates, especially ranefs; but think how to setup otherwise
                   iter=60,warmup=30,chains=2,
                   prior=c(
                       set_prior("normal(0,1.5)",class="b",nlpar=c("logitK","logitF")),
                       set_prior("normal(1,1)",class="b",nlpar=c("logr")),
                       set_prior("normal(0,1)",class="sd",nlpar=c("logitK","logitF","logr","residuals","obsprocess")),
                       set_prior("normal(0,1)",class="b",nlpar="invphi",lb=0),
                       set_prior("lkj(2)",class="cor")
                   ),
                   control=list(adapt_delta=0.9),seed=42)

##distance to the nearest patch and/or size of the nearest patch to M0??
## keep the last gen in the dataset (borrowing info good enough for trend?)
```

```{r}
ppc_ribbon(yrep=predict(mod_popsize,summary=FALSE),
           x=rank(predict(mod_popsize)[,1]),
           y=data_popsize$Peggs_est,
           prob = 0.5, prob_outer=0.95)
```


we need to keep the obs~treatment random effect as it reflects temporal stochasticity
refe: high stochasticity, low densities occasionally found even in core
restr: higher mean and low stochast, low densities rarely encountered in core

spatiotemp variab select for dispersal, spatio only against
helps interpret phenotype evolution, especially dispersal evolution in dens exp

restric core: reduced disp at densities effectively encountered in core
to check but at low densities, no change : lead to negDDD in core
# Body size

```{r}
data_size<-raw_size %>%
                mutate(IDgroup=ifelse(Generation==0,
                                      paste("Mix",Mix,"_stock",sep=""),
                                      paste("Mix",Mix,"_",Treatment,"_Rep",Replicate,"_GenerationFinal",sep="")
                                      )
                       )%>% 
                mutate(IDpop=ifelse(Generation==0,
                                    IDgroup,
                                    paste(IDgroup,"_Location",Location,sep=""))) %>% 
                mutate(IDindiv=paste(IDpop,ID_in_batch,sep="_")) %>% 
                mutate(trt=paste(Treatment,Location)) %>% 
                mutate(trt=relevel(factor(trt),"stock stock")) %>%
                mutate(trt=fct_recode(trt,stock="stock stock")) %>% 
                pivot_longer(cols=c("tibia_obsA","tibia_obsC"),names_to="observer",values_to="tibia") %>% 
                mutate(is.obsA=-0.5+1*(observer=="tibia_obsA"))

```

```{r}
### size accuracy
#
#mod=raw_size %>%
#                mutate(IDgroup=ifelse(Generation==0,
#                                      paste("Mix",Mix,"_stock",sep=""),
#                                      paste("Mix",Mix,"_",Treatment,"_Rep",Replicate,"_GenerationFinal",sep="")
#                                      )
#                       )%>% 
#                mutate(IDpop=ifelse(Generation==0,
#                                    IDgroup,
#                                    paste(IDgroup,"_Location",Location,sep=""))) %>% 
#                mutate(IDindiv=paste(IDpop,ID_in_batch,sep="_")) %>% 
#                mutate(trt=paste(Treatment,Location)) %>% 
#                mutate(trt=relevel(factor(trt),"stock stock")) %>%
#                mutate(trt=fct_recode(trt,stock="stock stock")) %>%
#  brm(mvbf(bf(tibia_obsC~1),bf(tibia_obsA~1),rescor=TRUE),
#      data=.
#)
```


use of nested hieracrhic groups to reflect phylogenetic relationships among pop taken from 

```{r}
mod_size=brm(bf(scale(tibia)~0+trt+is.obsA+(1|Mix/IDgroup/IDpop)+(1|gr(IDindiv,by=trt))),
        data=data_size,
        chains=4,iter=2000,
        prior=c(set_prior("normal(0,1)",class="b"),
                set_prior("normal(0,1)",class="sigma"),
                set_prior("normal(0,1)",class="sd")),
        seed=42
)

```

shifts mostly in variance. Looks negatively correlated with population density on aevrage in each category (though no way to be sure without stock densities)
read on effect of competition on size both plastic and selective


# Activity

```{r}
data_mvt<-raw_mvt %>% 
                filter(Generation %in% c(0,10)) %>% 
                filter(timestamp<=300) %>% ##some replicate are filmed longer than the nominal 5 minutes, this removes their ends to standardize
                mutate(Nobs=substr(File,nchar(File)-1,nchar(File))) %>% #note: rerecord correctly #wrong for G4 for which not given and last code is testing unit
                mutate(Nobs=as.numeric(str_remove(Nobs,"_"))) %>%  ##but we don't do G4 anyway, soooo
                mutate(IDgroup=ifelse(Generation==0,
                                      paste("Mix",Mix,"_stock",sep=""),
                                      paste("Mix",Mix,"_",Treatment,"_Rep",Replicate,"_GenerationFinal",sep="")
                                      )
                       )%>% 
                mutate(IDpop=ifelse(Generation==0,
                                    IDgroup,
                                    paste(IDgroup,"_Location",Location,sep=""))) %>% 
                mutate(IDtest=paste(IDpop,testing_unit,sep="_")) %>% 
                mutate(trt=paste(Treatment,Location)) %>% 
                mutate(trt=relevel(factor(trt),"stock stock")) %>%
                mutate(trt=fct_recode(trt,stock="stock stock")) %>% 
                group_by(Mix,Treatment,Generation,Replicate,Location,IDgroup,IDpop,IDtest,trt) %>% 
                summarise(percentData=mean(percentData), ##this column is already a testing_unit- level variable
                          activity_time=sum(path_duration*activity_percent),
                          obs_time=sum(path_duration),
                          Nobs=mean(Nobs)
                          ) %>% 
                mutate(activity_percent=activity_time/obs_time) %>% 
  ungroup()
```


```{r}
mod_mvt=brm(bf(activity_percent~0+trt+(1|Mix/IDgroup/IDpop),
               nlf(phi~10^log10phi),
               log10phi~percentData),
        data=data_mvt,family=Beta,
        chains=4,iter=2000,
        prior=c(set_prior("normal(0,1.5)",class="b"),
                set_prior("normal(0,1)",class="sd"),
                set_prior("normal(0,1)",class="b",nlpar="log10phi")),
        seed=42
)

```


# Fecundity 1

```{r}
data_fec<-raw_fec %>% 
                filter(Generation %in% c(0,12)) %>%
                mutate(IDgroup=ifelse(Generation==0,
                                      paste("Mix",Mix,"_stock",sep=""),
                                      paste("Mix",Mix,"_",Treatment,"_Rep",Replicate,"_GenerationFinal",sep="")
                                      )
                       )%>% 
                mutate(IDpop=ifelse(Generation==0,
                                    IDgroup,
                                    paste(IDgroup,"_Location",Location,sep=""))) %>% 
                mutate(trt=paste(Treatment,Location)) %>% 
                mutate(trt=relevel(factor(trt),"stock stock")) %>%
                mutate(trt=fct_recode(trt,stock="stock stock"))
```


```{r}
mod_fec=brm(bf(Fecundity~0+trt+(1|Mix/IDgroup/IDpop),
            zi~0+trt+(1|Mix/IDgroup/IDpop),
            nlf(shape~1/invshape),
            invshape~1),
         data=data_fec,family=zero_inflated_negbinomial(link_shape="identity"), #or hurdle?? think about it
         chains=4,iter=2000,
         prior=c(set_prior("normal(0,1.5)",class="b"),
                 set_prior("normal(0,1)",class="sd"),
                 set_prior("normal(0,1.5)",class="b",dpar="zi"),
                 set_prior("normal(0,1)",class="sd",dpar="zi"),
                 set_prior("normal(0,1)",nlpar="invshape",lb=0))
)
```


```{r}
### more accurate wrt protocol ## do the same for the DD version
data_fec$IDobs=1:length(data_fec$Mix)

mod_fec=brm(bf(Fecundity|trials(90)~0+trt+(1|Mix/IDgroup/IDpop)+(1|IDobs),
            zi~0+trt+(1|Mix/IDgroup/IDpop)),
         data=data_fec,family=zero_inflated_binomial(), #or hurdle?? think about it
         chains=4,iter=2000,
         prior=c(set_prior("normal(0,1.5)",class="b"),
                 set_prior("normal(0,1)",class="sd"),
                 set_prior("normal(0,1.5)",class="b",dpar="zi"),
                 set_prior("normal(0,1)",class="sd",dpar="zi"))
)
```



# Dispersal 1
```{r}
data_disp<-raw_disp %>% 
                filter(Generation %in% c(0,12)) %>%
                mutate(IDgroup=ifelse(Generation==0,
                                      paste("Mix",Mix,"_stock",sep=""),
                                      paste("Mix",Mix,"_",Treatment,"_Rep",Replicate,"_GenerationFinal",sep="")
                                      )
                       )%>% 
                mutate(IDpop=ifelse(Generation==0,
                                    IDgroup,
                                    paste(IDgroup,"_Location",Location,sep=""))) %>% 
                mutate(trt=paste(Treatment,Location)) %>% 
                mutate(trt=relevel(factor(trt),"stock stock")) %>%
                mutate(trt=fct_recode(trt,stock="stock stock")) %>% 
                mutate(Neggs_all=Neggs_start+Neggs_arrival) %>% 
                mutate(Neggs_all_scaled=scale(Neggs_all))
```


```{r}
mod_disp=brm(bf(Neggs_arrival|trials(Neggs_all)~0+trt+Neggs_all_scaled+(1|Mix/IDgroup/IDpop)),
         data=data_disp,family=binomial,
         chains=4,iter=200,
         prior=c(set_prior("normal(0,1.5)",class="b"),
                 set_prior("normal(0,1)",class="sd"))
)
```


loo/kfold to compare to version with no Neggs_scaled? seems to work (both here and density) but need to recheck because pareto bad obs

```{r}
mod_disp_fec=brm(bf(Neggs_all~0+trt+(1|Mix/IDgroup/IDpop)),
         data=data_disp,family=negbinomial,
         chains=4,iter=200,
         prior=c(set_prior("normal(0,1.5)",class="b"),
                 set_prior("normal(0,1)",class="sd"))
)
```


# Fecundity 2
```{r}
data_fec_dens<-raw_fec_dens %>% 
                mutate(IDgroup=ifelse(Generation==0,
                                      paste("Mix",Mix,"_stock",sep=""),
                                      paste("Mix",Mix,"_",Treatment,"_Rep",Replicate,"_GenerationFinal",sep="")
                                      )
                       )%>% 
                mutate(IDpop=ifelse(Generation==0,
                                    IDgroup,
                                    paste(IDgroup,"_Location",Location,sep=""))) %>% 
                mutate(trt=paste(Treatment,Location)) %>% 
                mutate(trt=relevel(factor(trt),"stock stock")) %>%
                mutate(trt=fct_recode(trt,stock="stock stock"))%>% 
                mutate(Density_centred=-0.5+1*(Density=="high"))
```


```{r}
mod_fec_dens=brm(bf(Fecundity~0+trt+trt:Density_centred+(1|Mix/IDgroup/IDpop),
            zi~0+trt+trt:Density_centred+(1|Mix/IDgroup/IDpop),
            nlf(shape~1/invshape),invshape~1),
         data=data_fec_dens,family=zero_inflated_negbinomial(link_shape = "identity"),
         chains=2,iter=200,
         prior=c(set_prior("normal(0,1.5)",class="b"),
                 set_prior("normal(0,1)",class="sd"),
                 set_prior("normal(0,1.5)",class="b",dpar="zi"),
                 set_prior("normal(0,1)",class="sd",dpar="zi"),
                 set_prior("normal(0,1)",nlpar="invshape",lb=0))
)
```
# Dispersal 2

```{r}
data_disp_dens<-raw_disp_dens %>% 
                mutate(IDgroup=ifelse(Generation==0,
                                      paste("Mix",Mix,"_stock",sep=""),
                                      paste("Mix",Mix,"_",Treatment,"_Rep",Replicate,"_GenerationFinal",sep="")
                                      )
                       )%>% 
                mutate(IDpop=ifelse(Generation==0,
                                    IDgroup,
                                    paste(IDgroup,"_Location",Location,sep=""))) %>% 
                mutate(trt=paste(Treatment,Location)) %>% 
                mutate(trt=relevel(factor(trt),"stock stock")) %>%
                mutate(trt=fct_recode(trt,stock="stock stock")) %>% 
                mutate(Neggs_all=Neggs_start+Neggs_arrival) %>% 
                mutate(Neggs_all_scaled=scale(Neggs_all)) %>% 
                mutate(Neggs_all_log=log(Neggs_all)) %>% 
                mutate(Density_centred=-0.5+1*(Density=="high"))
```


```{r}
mod_disp_dens=brm(bf(Neggs_arrival|trials(Neggs_all)~0+trt+trt:Density_centred+Neggs_all_scaled+(1|Mix/IDgroup/IDpop)),
         data=data_disp_dens,family=binomial,
         chains=4,iter=200,
         prior=c(set_prior("normal(0,1.5)",class="b"),
                 set_prior("normal(0,1)",class="sd"))
)
```


```{r summary-model}
###prediction intervals around each point
ppc_ribbon(yrep=(predict(mod_disp_dens,summary=FALSE)),
           x=rank(predict(mod_disp_dens)[,1]),
           y=data_disp_dens$Neggs_arrival,
           prob = 0.5, prob_outer=0.95)
```


```{r}
mod_disp_dens_fec=brm(bf(Neggs_all~0+trt*Density_centred+(1|Mix/IDgroup/IDpop)),
         data=data_disp_dens,family=negbinomial,
         chains=4,iter=200,
         prior=c(set_prior("normal(0,1.5)",class="b"),
                 set_prior("normal(0,1)",class="sd"))
)
```