---
title: "Code for: \"Landscape connectivity alters the evolution of density-dependent dispersal during pushed range expansions\" "
author: "Maxime Dahirel, Aline Bertin, Vincent Calcagno, Camille Duraj, Simon Fellous, Géraldine Groussier, Ludovic Mailleret, Anaël Marchand, Elodie Vercken (this code by M. Dahirel)"
date:
output: 
  html_document:
    theme: yeti
    toc: TRUE
    toc_float: TRUE
    code_download: TRUE
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(arm)
library(tidyverse)

library(cmdstanr) 
#library(rstan) 
#rstan can be used as alternate backend; just don't forget to change the backend argument in each model
library(brms)
options(mc.cores = 2)

library(bayesplot)

library(tidybayes)

library(ggnewscale)
library(patchwork)

library(here)
```

# Context


#load data

```{r}
raw_size <- read_csv(here("data","exp1_bodysize.csv"))

raw_mvt <- read_csv(here("data","exp2_movement.csv"))

raw_fec <- read_csv(here("data","exp2_fecundity.csv"))
raw_disp <- read_csv(here("data","exp2_dispersal.csv"))

raw_fec_dens <- read_csv(here("data","exp3_fecundity.csv"))
raw_disp_dens <- read_csv(here("data","exp3_dispersal.csv"))
```

data were collected at some intermediate times in some experiments. For this paper, we only use experiment endpoints

# Body size

ID group is the landscape ID (so Mix * Treatment * Replicate); = Mix "stock" if stock
we recode Treatment and Location levels
then IDpop is the population replicate, = ID group for stock, = IDgroup*Location (edge-core) for others
then IDindividual for which individual is measured
then creation of a context variable (basically treatment*location)
then pivoting to long format, so one row = one obs instead of one individual

```{r}
data_size<-raw_size %>%
                mutate(IDgroup=ifelse(Generation==0,
                                      paste("Mix",Mix,"_stock",sep=""),
                                      paste("Mix",Mix,"_",Treatment,"_Rep",Replicate,"_GenerationFinal",sep="")
                                      )
                       )%>% 
                mutate(Location=fct_recode(factor(Location),edge="front")) %>% 
                mutate(Treatment=fct_recode(factor(Treatment),
                                            reference="control",
                                            `reduced connectivity`="restricted connectedness")) %>% 
                mutate(IDpop=ifelse(Generation==0,
                                    IDgroup,
                                    paste(IDgroup,"_Location",Location,sep=""))) %>% 
                mutate(IDindiv=paste(IDpop,ID_in_batch,sep="_")) %>% 
                mutate(context=paste(Treatment,Location)) %>% 
                mutate(context=fct_recode(context,stock="stock stock")) %>% 
                mutate(context = fct_relevel(context,"stock", after=0)) %>%
                pivot_longer(cols=c("tibia_obsA","tibia_obsC"),names_to="observer",values_to="tibia") %>% 
  filter(is.na(tibia)==FALSE)

```


use of nested hieracrhic groups to reflect phylogenetic relationships among pop taken from 

```{r}
if (file.exists(here("R_output", "model_size.Rdata")))
# this if-else statement is avoid re-fitting a model when knitting Rmd file if there is already one existing in R_output
# to override, re-run the model and re-save manually by selecting relevant code lines then knit (or delete the Rdata object)
  {
    load(here("R_output", "model_size.Rdata"))
  } else {


mod_size=brm(bf(scale(tibia)~context+(1|Mix/IDgroup/IDpop)+(1|IDindiv)),
        data=data_size,
        chains=4,iter=4000,
        prior=c(set_prior("normal(0,1)",class="Intercept"),
                set_prior("normal(0,1)",class="b"),
                set_prior("normal(0,1)",class="sigma"),
                set_prior("normal(0,1)",class="sd")),
        seed=42,control=list(adapt_delta=0.99,max_treedepth=15), backend="cmdstanr"
)

  save(list = "mod_size", file = here("R_output", "model_size.Rdata"))
}
```


```{r}
preds_size <- data_size %>% 
  select(context,Location,Treatment) %>% 
  mutate(Treatment=fct_relevel(factor(Treatment),"stock",after=0)) %>% 
  unique() %>% 
  add_fitted_draws(mod_size,re_formula = NA) %>% 
  ungroup() %>% 
  mutate(.value=.value* (sd(data_size$tibia,na.rm=TRUE))+ (mean(data_size$tibia,na.rm=TRUE))) %>% 
  select(.draw,.value,Treatment,Location,context)

ggplot(preds_size) +
  geom_jitter(data=data_size %>% 
                group_by(Location,Treatment,IDindiv) %>% 
                summarise(tibia=mean(tibia)),aes(x=Location,y=tibia),col="grey50",alpha=0.4)+
  stat_eye(aes(x=Location,y=.value),.width=c(0.01,0.95), slab_alpha=0.7)+
  scale_x_discrete("Experimental context")+
  scale_y_continuous("Mean tibia length (μm)")+
  facet_grid(cols=vars(Treatment),scales="free_x",space="free_x")+
  theme_bw()

```


# Activity

```{r}
data_mvt<-raw_mvt %>% 
                filter(Generation %in% c(0,10)) %>% 
                filter(timestamp<=300) %>% ##some replicate are filmed slightly longer than the nominal 5 minutes, this removes their ends to standardize
                mutate(IDgroup=ifelse(Generation==0,
                                      paste("Mix",Mix,"_stock",sep=""),
                                      paste("Mix",Mix,"_",Treatment,"_Rep",Replicate,"_Generation10",sep="")
                                      )
                       )%>% 
                mutate(Location=fct_recode(factor(Location),edge="front")) %>% 
                mutate(Treatment=fct_recode(factor(Treatment),
                                            reference="control",
                                            `reduced connectivity`="restricted connectedness")) %>% 
                mutate(IDpop=ifelse(Generation==0,
                                    IDgroup,
                                    paste(IDgroup,"_Location",Location,sep=""))) %>% 
                mutate(IDtest=paste(IDpop,testing_unit,sep="_")) %>% 
                mutate(context=paste(Treatment,Location)) %>% 
                mutate(context=fct_recode(context,stock="stock stock")) %>% 
                mutate(context = fct_relevel(context,"stock", after=0)) %>%
                group_by(Mix,Treatment,Generation,Replicate,Location,IDgroup,IDpop,IDtest,context) %>% 
                summarise(percentData=mean(percentData), ##this column is already a testing_unit- level variable
                          activity_time=sum(path_duration*activity_percent),
                          obs_time=sum(path_duration)
                          ) %>% 
                mutate(activity_percent=activity_time/obs_time) %>% 
  ungroup()
```


```{r}
if (file.exists(here("R_output", "model_mvt.Rdata")))
# this if-else statement is avoid re-fitting a model when knitting Rmd file if there is already one existing in R_output
# to override, re-run the model and re-save manually by selecting relevant code lines then knit (or delete the Rdata object)
  {
    load(here("R_output", "model_mvt.Rdata"))
  } else {
mod_mvt=brm(bf(activity_percent~context+(1|Mix/IDgroup/IDpop),
               nlf(phi~1/invphi),
               invphi~1),
        data=data_mvt,family=Beta,
        chains=4,iter=6000,
        prior=c(set_prior("normal(0,1.5)",class="Intercept"),
                set_prior("normal(0,1)",class="b"),
                set_prior("normal(0,1)",class="sd"),
                set_prior("normal(0,1)",nlpar="invphi",lb=0)),
        seed=42,control=list(adapt_delta=0.99),backend="cmdstanr"
)
  save(list = "mod_mvt", file = here("R_output", "model_mvt.Rdata"))
}
```


```{r}
preds_mvt <- data_mvt %>% 
  select(context,Location,Treatment) %>% 
  mutate(Treatment=fct_relevel(factor(Treatment),"stock",after=0)) %>% 
  distinct() %>% 
  add_fitted_draws(mod_mvt,re_formula = NA) %>% 
  ungroup()

ggplot(preds_mvt) +
  geom_jitter(data=data_mvt,aes(x=Location,y=activity_percent, size=obs_time),col="grey50",alpha=0.4)+
  scale_size(range=c(1,4), guide=FALSE)+
  ggnewscale::new_scale("size")+
  stat_eye(aes(x=Location,y=.value), .width=c(0.01,0.95), slab_alpha=0.7)+
  scale_x_discrete("Experimental context")+
  scale_y_continuous("Proportion of time active")+
  facet_grid(cols=vars(Treatment),scales="free_x",space="free_x")+
  theme_bw()

#posterior grand mean
preds_mvt %>% 
  group_by(.draw) %>% 
  summarise(grandmean=mean(.value)) %>% 
  mean_hdi(grandmean)

```


# Fecundity 1

```{r}
data_fec<-raw_fec %>% 
                filter(Generation %in% c(0,12)) %>%
                mutate(IDgroup=ifelse(Generation==0,
                                      paste("Mix",Mix,"_stock",sep=""),
                                      paste("Mix",Mix,"_",Treatment,"_Rep",Replicate,"_GenerationFinal",sep="")
                                      )
                       )%>% 
                mutate(Location=fct_recode(factor(Location),edge="front")) %>% 
                mutate(Treatment=fct_recode(factor(Treatment),
                                            reference="control",
                                            `reduced connectivity`="restricted connectedness")) %>%
                mutate(IDpop=ifelse(Generation==0,
                                    IDgroup,
                                    paste(IDgroup,"_Location",Location,sep=""))) %>% 
                mutate(context=paste(Treatment,Location)) %>% 
                mutate(context=fct_recode(context,stock="stock stock"))%>%
                mutate(context = fct_relevel(context,"stock", after=0)) 

data_fec$obsID=1:dim(data_fec)[1]
```


```{r}
if (file.exists(here("R_output", "model_fecundity1.Rdata")))
# this if-else statement is avoid re-fitting a model when knitting Rmd file if there is already one existing in R_output
# to override, re-run the model and re-save manually by selecting relevant code lines then knit (or delete the Rdata object)
  {
    load(here("R_output", "model_fecundity1.Rdata"))
  } else {
mod_fec=brm(bf(Fecundity~context+(1|Mix/IDgroup/IDpop),
            zi~context+(1|Mix/IDgroup/IDpop),
            nlf(shape~1/invshape),
               invshape~1),
         data=data_fec,family=zero_inflated_negbinomial(),
         chains=4,iter=4000,
         prior=c(set_prior("normal(3.8,0.5)",class="Intercept"),
                 set_prior("normal(0,1)",class="b"),
                 set_prior("normal(0,1)",class="sd"),
                 set_prior("normal(0,0.2)",class="sd",group="Mix"), ## narrower prior for mix to stamp divergences
                 set_prior("normal(0,1.5)",class="Intercept",dpar="zi"),
                 set_prior("normal(0,1)",class="b",dpar="zi"),
                 set_prior("normal(0,1)",class="sd",dpar="zi"),
                 set_prior("normal(0,0.2)",class="sd",dpar="zi",group="Mix"),
                set_prior("normal(0,1)",nlpar="invshape",lb=0)),
         seed=42,control=list(adapt_delta=0.99,max_treedepth=20),backend="cmdstanr"
)

  save(list = "mod_fec", file = here("R_output", "model_fecundity1.Rdata"))
  }

```




# Fecundity 2
```{r}
data_fec_dens<-raw_fec_dens %>% 
                mutate(IDgroup=ifelse(Generation==0,
                                      paste("Mix",Mix,"_stock",sep=""),
                                      paste("Mix",Mix,"_",Treatment,"_Rep",Replicate,"_GenerationFinal",sep="")
                                      )
                       )%>% 
                mutate(Location=fct_recode(factor(Location),edge="front")) %>% 
                mutate(Treatment=fct_recode(factor(Treatment),
                                            reference="control",
                                            `reduced connectivity`="restricted connectedness")) %>%
                mutate(IDpop=ifelse(Generation==0,
                                    IDgroup,
                                    paste(IDgroup,"_Location",Location,sep=""))) %>% 
                mutate(context=paste(Treatment,Location)) %>% 
                mutate(context=fct_recode(context,stock="stock stock"))%>% 
                mutate(context = fct_relevel(context,"stock", after=0)) %>% 
                mutate(Density=fct_relevel(factor(Density),"high",after=Inf)) %>% 
                mutate(Density_centred = as.numeric(Density=="high") - mean(as.numeric(Density=="high")))

data_fec_dens$obsID=1:dim(data_fec_dens)[1]
```

```{r}
if (file.exists(here("R_output", "model_fecundity2.Rdata")))
# this if-else statement is avoid re-fitting a model when knitting Rmd file if there is already one existing in R_output
# to override, re-run the model and re-save manually by selecting relevant code lines then knit (or delete the Rdata object)
  {
    load(here("R_output", "model_fecundity2.Rdata"))
  } else {
mod_fec_dens=brm(bf(Fecundity~context*Density_centred+(1|Mix/IDgroup/IDpop),
            zi~context*Density_centred+(1|Mix/IDgroup/IDpop),
            nlf(shape~1/invshape),
               invshape~1),
         data=data_fec_dens,family=zero_inflated_negbinomial,
         chains=4,iter=4000,
         prior=c(set_prior("normal(3.8,0.5)",class="Intercept"),
                 set_prior("normal(0,1)",class="b"),
                 set_prior("normal(0,1)",class="sd"),
                 set_prior("normal(0,0.2)",class="sd",group="Mix"), ## narrower prior for mix to stamp divergences
                 set_prior("normal(0,1.5)",class="Intercept",dpar="zi"),
                 set_prior("normal(0,1)",class="b",dpar="zi"),
                 set_prior("normal(0,1)",class="sd",dpar="zi"),
                 set_prior("normal(0,0.2)",class="sd",dpar="zi",group="Mix"),
                set_prior("normal(0,1)",nlpar="invshape",lb=0)),
         seed=42,control=list(adapt_delta=0.99,max_treedepth=20),backend="cmdstanr"
)
  save(list = "mod_fec_dens", file = here("R_output", "model_fecundity2.Rdata"))
}
```



```{r summary-model}
###prediction intervals around each point
ppc_ribbon(yrep=(predict(mod_fec_dens,summary=FALSE)),
           x=rank(predict(mod_fec_dens)[,1]),
           y=data_fec_dens$Fecundity,
           prob = 0.5, prob_outer=0.95)
```

```{r}

p1 <- data_fec %>% 
  select(Location, Treatment, context) %>% 
  distinct() %>% 
  add_fitted_draws(mod_fec,re_formula=NA) %>% 
  mutate(Treatment = fct_relevel(Treatment,"reference", after=Inf)) %>% 
  mutate(Treatment = fct_relevel(Treatment,"reduced connectivity", after=Inf)) %>% 
  ggplot()+
  geom_jitter(data=data_fec,
              aes(Location,Fecundity),
              col="#d95f02",alpha=0.4)+
  stat_eye(aes(Location, .value),.width=c(0.01,0.95),fill="#d95f02", slab_alpha=0.7)+
  scale_x_discrete("")+
  scale_y_continuous("# of hosts parasitised")+
  facet_grid(cols=vars(Treatment),scales="free_x",space="free_x")+
  coord_cartesian(ylim=c(0,90))+
  theme_bw()

p2 <- data_fec_dens %>%  
  select(Location, Treatment, context, Density, Density_centred) %>% 
  distinct() %>% 
  add_fitted_draws(mod_fec_dens,re_formula=NA) %>%
  ungroup() %>% 
  mutate(Treatment = fct_relevel(Treatment,"reference", after=Inf)) %>% 
  mutate(Treatment = fct_relevel(Treatment,"reduced connectivity", after=Inf)) %>% 
  ggplot()+
  geom_point(data=data_fec_dens,
             aes(Location,Fecundity, col=Density),
             position=position_jitterdodge(dodge.width=1), alpha=0.4)+
  stat_eye(aes(Location, .value,fill=Density), .width=c(0.01,0.95), position="dodge", slab_alpha=0.7)+
  scale_x_discrete("")+
  scale_y_continuous("# of hosts parasitised")+
  scale_fill_manual(values=c("#d95f02","#7570b3")) + 
  scale_colour_manual(values=c("#d95f02","#7570b3")) + 
  facet_grid(cols=vars(Treatment),scales="free_x",space="free_x")+
  coord_cartesian(ylim=c(0,90))+
  theme_bw()+
  guides(size = FALSE)

p3 <- data_fec_dens %>%  
  select(Location, Treatment, context, Density, Density_centred) %>% 
  distinct() %>% 
  add_fitted_draws(mod_fec_dens,re_formula=NA) %>%
  mutate(Treatment = fct_relevel(Treatment,"reference", after=Inf)) %>% 
  mutate(Treatment = fct_relevel(Treatment,"reduced connectivity", after=Inf)) %>% 
  ungroup() %>% 
  select(Location, Treatment, Density, .value, .draw) %>% 
  pivot_wider(names_from=Density,values_from=.value) %>% 
  mutate(diff=log(high)-log(low)) %>% 
  ggplot()+
  stat_eye(aes(Location, diff), .width=c(0.01,0.95),slab_alpha=0.7)+
  geom_hline(yintercept=0,lty=2)+
  scale_x_discrete("Experimental context")+
  scale_y_continuous("Net effect of density (log scale)")+
  facet_grid(cols=vars(Treatment),scales="free_x",space="free_x")+
  coord_cartesian(ylim=c(-1,1))+
  theme_bw()

(p1/p2/p3) & plot_annotation(tag_levels = "A")

```


# Dispersal 1
```{r}
data_disp<-raw_disp %>% 
                filter(Generation %in% c(0,12)) %>%
                mutate(IDgroup=ifelse(Generation==0,
                                      paste("Mix",Mix,"_stock",sep=""),
                                      paste("Mix",Mix,"_",Treatment,"_Rep",Replicate,"_GenerationFinal",sep="")
                                      )
                       )%>% 
                mutate(Location=fct_recode(factor(Location),edge="front")) %>% 
                mutate(Treatment=fct_recode(factor(Treatment),
                                            reference="control",
                                            `reduced connectivity`="restricted connectedness")) %>%
                mutate(IDpop=ifelse(Generation==0,
                                    IDgroup,
                                    paste(IDgroup,"_Location",Location,sep=""))) %>% 
                mutate(context=paste(Treatment,Location)) %>% 
                mutate(context=relevel(factor(context),"stock stock")) %>%
                mutate(context=fct_recode(context,stock="stock stock")) %>% 
                mutate(context = fct_relevel(context,"stock", after=0)) %>%
                mutate(Neggs_all=Neggs_start+Neggs_arrival) %>% 
                mutate(Neggs_all_scaled=scale(Neggs_all))
```


```{r}
if (file.exists(here("R_output", "model_dispersal1.Rdata")))
# this if-else statement is avoid re-fitting a model when knitting Rmd file if there is already one existing in R_output
# to override, re-run the model and re-save manually by selecting relevant code lines then knit (or delete the Rdata object)
  {
    load(here("R_output", "model_dispersal1.Rdata"))
  } else {
mod_disp=brm(bf(Neggs_arrival|trials(Neggs_all)~context+Neggs_all_scaled+(1|Mix/IDgroup/IDpop)),
         data=data_disp,family=binomial,
         chains=4,iter=4000,
         prior=c(set_prior("normal(0,1.5)",class="Intercept"),
                 set_prior("normal(0,1)",class="b"),
                 set_prior("normal(0,1)",class="sd")),
         seed=42,control=list(adapt_delta=0.99), backend="cmdstanr"
)
  save(list = "mod_disp", file = here("R_output", "model_dispersal1.Rdata"))
}
```

think about adding the model minus Neggs_all_scaled and compare

```{r}
if (file.exists(here("R_output", "model_dispersal1_alt.Rdata")))
# this if-else statement is avoid re-fitting a model when knitting Rmd file if there is already one existing in R_output
# to override, re-run the model and re-save manually by selecting relevant code lines then knit (or delete the Rdata object)
  {
    load(here("R_output", "model_dispersal1_alt.Rdata"))
  } else {
mod_disp_alt=brm(bf(Neggs_arrival|trials(Neggs_all)~context+(1|Mix/IDgroup/IDpop)),
         data=data_disp,family=binomial,
         chains=4,iter=4000,
         prior=c(set_prior("normal(0,1.5)",class="Intercept"),
                 set_prior("normal(0,1)",class="b"),
                 set_prior("normal(0,1)",class="sd")),
         seed=42,control=list(adapt_delta=0.99), backend="cmdstanr"
)
  save(list = "mod_disp_alt", file = here("R_output", "model_dispersal1_alt.Rdata"))
}
```


# Dispersal 2

```{r}

data_disp_dens<-raw_disp_dens %>% 
                mutate(IDgroup=ifelse(Generation==0,
                                      paste("Mix",Mix,"_stock",sep=""),
                                      paste("Mix",Mix,"_",Treatment,"_Rep",Replicate,"_GenerationFinal",sep="")
                                      )
                       )%>% 
                mutate(Location=fct_recode(factor(Location),edge="front")) %>% 
                mutate(Treatment=fct_recode(factor(Treatment),
                                            reference="control",
                                            `reduced connectivity`="restricted connectedness")) %>%
                mutate(IDpop=ifelse(Generation==0,
                                    IDgroup,
                                    paste(IDgroup,"_Location",Location,sep=""))) %>% 
                mutate(context=paste(Treatment,Location)) %>% 
                mutate(context=relevel(factor(context),"stock stock")) %>%
                mutate(context=fct_recode(context,stock="stock stock")) %>% 
                mutate(context = fct_relevel(context,"stock", after=0)) %>%
                mutate(Neggs_all=Neggs_start+Neggs_arrival) %>% 
                mutate(Neggs_all_scaled=scale(Neggs_all)) %>% 
                mutate(Density=fct_relevel(factor(Density),"high",after=Inf)) %>% 
                mutate(Density_centred = as.numeric(Density=="high") - mean(as.numeric(Density=="high")))
```


```{r}
if (file.exists(here("R_output", "model_dispersal2.Rdata")))
# this if-else statement is avoid re-fitting a model when knitting Rmd file if there is already one existing in R_output
# to override, re-run the model and re-save manually by selecting relevant code lines then knit (or delete the Rdata object)
  {
    load(here("R_output", "model_dispersal2.Rdata"))
  } else {
mod_disp_dens=brm(bf(Neggs_arrival|trials(Neggs_all)~context*Density_centred+Neggs_all_scaled+(1|Mix/IDgroup/IDpop)),
         data=data_disp_dens,family=binomial,
         chains=4,iter=4000,
         prior=c(set_prior("normal(0,1.5)",class="Intercept"),
                 set_prior("normal(0,1)",class="b"),
                 set_prior("normal(0,1)",class="sd")),
         seed=42,control=list(adapt_delta=0.99),backend="cmdstanr"
)

  save(list = "mod_disp_dens", file = here("R_output", "model_dispersal2.Rdata"))
}
```

```{r}
if (file.exists(here("R_output", "model_dispersal2_alt.Rdata")))
# this if-else statement is avoid re-fitting a model when knitting Rmd file if there is already one existing in R_output
# to override, re-run the model and re-save manually by selecting relevant code lines then knit (or delete the Rdata object)
  {
    load(here("R_output", "model_dispersal2_alt.Rdata"))
  } else {
mod_disp_dens_alt=brm(bf(Neggs_arrival|trials(Neggs_all)~context*Density_centred+(1|Mix/IDgroup/IDpop)),
         data=data_disp_dens,family=binomial,
         chains=4,iter=4000,
         prior=c(set_prior("normal(0,1.5)",class="Intercept"),
                 set_prior("normal(0,1)",class="b"),
                 set_prior("normal(0,1)",class="sd")),
         seed=42,control=list(adapt_delta=0.99),backend="cmdstanr"
)

  save(list = "mod_disp_dens_alt", file = here("R_output", "model_dispersal2_alt.Rdata"))
}
```

```{r summary-model}
###prediction intervals around each point
ppc_ribbon(yrep=(predict(mod_disp_dens,summary=FALSE)),
           x=rank(predict(mod_disp_dens)[,1]),
           y=data_disp_dens$Neggs_arrival,
           prob = 0.5, prob_outer=0.95)
```


```{r}

p1 <- data_disp %>% 
  mutate(Neggs_all_scaled = mean(Neggs_all_scaled),
         Neggs_all = 1) %>% 
  select(Location, Treatment, context, Neggs_all_scaled, Neggs_all) %>% 
  distinct() %>% 
  add_fitted_draws(mod_disp,re_formula=NA) %>% 
  mutate(Treatment = fct_relevel(Treatment,"reference", after=Inf)) %>% 
  mutate(Treatment = fct_relevel(Treatment,"reduced connectivity", after=Inf)) %>% 
  ggplot()+
  geom_jitter(data=data_disp,
              aes(Location,Neggs_arrival/Neggs_all, size=Neggs_all),
              col="#d95f02",alpha=0.4)+  
  scale_size(range=c(1,4), guide=FALSE)+
  ggnewscale::new_scale("size")+
  stat_eye(aes(Location, .value),.width=c(0.01,0.95),fill="#d95f02", slab_alpha=0.7)+
  scale_x_discrete("")+
  scale_y_continuous("Effective dispersal rate")+
  facet_grid(cols=vars(Treatment),scales="free_x",space="free_x")+
  coord_cartesian(ylim=c(0,1))+
  theme_bw()


p2 <- data_disp_dens %>%  
  mutate(Neggs_all_scaled = mean(Neggs_all_scaled),
         Neggs_all = 1) %>% 
  select(Location, Treatment, context, Neggs_all_scaled, Neggs_all, Density, Density_centred) %>% 
  distinct() %>% 
  add_fitted_draws(mod_disp_dens,re_formula=NA) %>% 
  mutate(Treatment = fct_relevel(Treatment,"reference", after=Inf)) %>% 
  mutate(Treatment = fct_relevel(Treatment,"reduced connectivity", after=Inf)) %>% 
  ggplot()+
  geom_point(data=data_disp_dens,
             aes(Location,Neggs_arrival/Neggs_all, col=Density, size=Neggs_all),
             position=position_jitterdodge(dodge.width=1), alpha=0.4)+
  scale_size(range=c(1,4), guide=FALSE)+
  ggnewscale::new_scale("size")+
  stat_eye(aes(Location, .value,fill=Density), .width=c(0.01,0.95), position="dodge", slab_alpha=0.7)+
  scale_x_discrete("")+
  scale_y_continuous("Effective dispersal rate")+
  scale_fill_manual(values=c("#d95f02","#7570b3")) + 
  scale_colour_manual(values=c("#d95f02","#7570b3")) + 
  facet_grid(cols=vars(Treatment),scales="free_x",space="free_x")+
  coord_cartesian(ylim=c(0,1))+
  theme_bw()

p3 <- data_disp_dens %>%  
  mutate(Neggs_all_scaled = mean(Neggs_all_scaled),
         Neggs_all = 1) %>% 
  select(Location, Treatment, context, Neggs_all_scaled, Neggs_all, Density, Density_centred) %>% 
  distinct() %>% 
  add_fitted_draws(mod_disp_dens,re_formula=NA) %>% 
  ungroup() %>% 
  mutate(Treatment = fct_relevel(Treatment,"reference", after=Inf)) %>% 
  mutate(Treatment = fct_relevel(Treatment,"reduced connectivity", after=Inf)) %>% select(Location, Treatment, Density, .value, .draw) %>% 
  pivot_wider(names_from=Density,values_from=.value) %>% 
  mutate(diff=logit(high)-logit(low)) %>% 
  ggplot()+
  stat_eye(aes(Location, diff), .width=c(0.01,0.95),slab_alpha=0.7)+
  geom_hline(yintercept=0,lty=2)+
  scale_x_discrete("Experimental context")+
  scale_y_continuous("Net effect of density (logit scale)")+
  facet_grid(cols=vars(Treatment),scales="free_x",space="free_x")+
  coord_cartesian(ylim=c(-3,3))+
  theme_bw()


(p1/p2/p3) & plot_annotation(tag_levels = "A")

```

 easy to replace the models with their alt in chunk above to see inferences remain especially relative to DDD
